Developing a preprint culture in biology

### 2012-10-18 ms.tex
First round of small editions

Public preprint servers allow authors to make manuscripts publicly available
before, or in parallel to, submitting them to journals for traditional peer-
review. The rationale for preprint servers is fundamentally simple: to make
the results of scientific research available to the scientific community as
soon as possible, instead of waiting until the peer-review process is fully
completed.  Sharing manuscripts using preprint servers has numerous advantages
including: 1) rapid dissemination of work-in-progress to a wider audience; 2)
immediate visibility of the research output for early-carreer scientsits; 3)
improved peer review by encouraging feedback from the entire research
community; and 4) a fair and straightforward way to establish precedence.

Open preprint servers offer a great opportunity for open science, especially if
the community embraces the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work \cite{lom12}. These initiatives are
important to fully exploit the potential of open preprints servers. Posting
preprints online increases the community of available informal peer reviewers,
and uses the internet for its original community-building purposes.

Preprints began to gain popularity 20 years ago with the advent of arXiv, an open
preprint server widely used in physics and mathematics \cite{gin11}. Preprints
are also integral to the culture of other scientific fields.  Paul Krugman noted
that, in economics, the \emph{traditional model of submit, get refereed,
publish, and then people will read your work broke down a long time ago. In
fact, it had more or less fallen apart by the early 80s} \cite{kru12}. In
addition to a section in arXiv, economists have also the RePEc (Research Papers
in Economics) initiative, which aims to create an archive of working papers,
manuscripts, and book chapters.

Despite the success of this approach in other fields, most manuscripts in
biology are not posted to preprint servers and are therefore not seen by more
than a handful of other scientists prior to publication. In this article, we: 1)
highlight the advantages of open preprint servers for both scientists and
publishers; 2) address several misconceptions about preprints that are common
among biologists; 3) discuss the preprint policies of major publishers in
biology; and 4) review the most popular preprint servers currently available.

@@

Public preprints can be crucial to early-career scientists. The delay before
publication is seldom compatible with the pressure to show an impressive
publication record when applying for a scholarship or a position. Increasing
the perceived value of pre-prints as close, or equal, to journal articles (a
point we discuss in the conclusion) will allow young researchers to put their
research outcome in the open, and build a reputation for themselves through
the diffusion of their work, without the fear that this work will not be
recognized by grant or job commitees. In this perspective, mixed publication
systems like experimented by F1000Research (see below) should be encouraged as
they combine the best of both worlds: contributions are initally considered
preprints, but acquire full article status after positive reviews are
received.

Posting manuscripts as preprints also has the potential to result in higher
quality science, by allowing for pre-publication feedback from a large pool of
reviewers. Prepublication reviews by a small network of colleagues are common in
the biological sciences and are an important part of the scientific process.
These ``friendly'' reviews increases the chances of errors being caught prior to
publication. Furthermore, the formal peer-review process as a whole is
critically over-loaded. As the number of active scientists increases and the
pressure to publish increases, it has become increasingly difficult for journals
to find reviewers \cite{hoc09}.  At the same time, rejection rates are high in
most journals \cite{aar08,roh09}, and when not invited to submit a revision,
authors must start the process over again at another journal. As a result,
initiatives to reduce time from submission to publication have emerged across
the scientific community. Rohr et al.  \cite{roh09} called for the recycling and
reuse of peer-reviews: by attaching previous reviews and detailed replies to a
new submission, both the editor and the referees can gauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. A widespread use of
preprint servers can achieve the same goal of reducing the time spent in review.
With a rich enough community of scientists depositing preprints, and commenting
on them, the process of an open pre-review can become widespread and will
overall increase the quality of first submissions \cite{hoc12}.

@@

Finally, public preprint servers offer a fair way to establish intellectual
priority by making the work available as soon as it is complete. Some
manuscripts will spend much more time in the review process and/or in
production after acceptance, than others. This means that publication and
acceptance dates do not accurately characterize who came up with an idea
first. For this reason, mathematicians and physicists have embraced arXiv in
part to establish priority in a fair way \cite{gin11,cal12}.

@@

There are a number of reasons that biologists have not traditionally had a
culture of sharing preprints, many of which are based on common misconceptions
regarding the costs and benefits associated with the practice. For example, in contrast
to other fields there is a perception in biology that public preprints
make it easier to steal ideas. However, there is no evidence of this happening
in the numerous other fields that have adopted preprint servers, and since preprint
servers create a clear record of who had the idea first, and when, this
appears to be a largely unfounded concern. In other fields preprints serve the
opposite role, they allow straightforward establishment of precedence, letting
research lay claim to an idea thus preventing it from being "stolen" \cite{gin11}

Another major concern is that if a researcher posts a public preprint that they
will not be allowed to submit their paper to their journal of choice.  This is
based on a certain interpretation of the Ingelfinger rule: scientists should not
publish the same manuscript twice \cite{alt96}. A preprint is simply a document
that allows ideas to spread and be discussed, it is not yet formally validated
by the peer-review system. This is why the majority of publishers do not see
arXiv and similar services as a violation of the Ingelfinger rule. Almost all of
the major publishers in biology are preprint-friendly, including: Nature
Publishing Group, PLOS, BMC, PNAS, Elsevier, and Springer \ref{table:policies}.
The Ecological Society of America and the Genetics Society of America recently
changed their policies to allow public preprints. In fact, \emph{Nature}
responded to the rumour that they refused manuscripts submitted to arXiv by
saying that ``\emph{Nature} never wishes to stand in the way of communication
between researchers. We seek rather to add value for authors and the community
at large in our peer review, selection and editing'' \cite{nat05}.  Still, a few
journals adopt a ``by default'' hostile attitude towards preprints, mostly due
to the lack of clear policy of the publishers, or perhaps because a preprint
culture has not developed in biology and the practice is still considered
unusual. As an example, Wiley-Blackwell, which publishes some of the leading
journals in biology, has no official policy on the matter.

@@

One of the advantages of figshare over arXiv for biologists is that is it not
limited to quantitative sciences. arXiv.org has sections on quantitative biology
but might not be appropriate for non-quantitative work. With its flexible
approach to preprints, figshare offers an alternative to arXiv for empirical
biologists. Furthermore, by allowing all types of content, figshare also
provides an archive for early results (e.g.: figures, lab presentations).

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new Open Access
publisher that combines both a preprint server, and a peer reviewed journal.  It
is focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a place in biology. Like figshare this is
an advantage relative to arXiv for biologists doing non-quantitative work.  Also
like figshare, PeerJ allows commenting on posted preprints, improving the
potential for pre-publication dialog. In addition, preprints can be made private
if the authors choose, and shared only with selected colleagues. While this
reduces many of the benefits of preprints described above, it may allow some
researchers who would not otherwise post preprints to begin to explore the
possibility in a manner appropriate to their current circumstances.

In contrast to other preprint servers users cannot post unlimited public
preprints for free. One preprint per year can be posted for free, with a onetime
(\emph{i.e.} lifetime) fee of 99 dollars required to allow an author to post unlimited
public preprints. The preprint server is not tied to the journal, so preprints
can be posted regardless of where they will eventually be submitted for
publication.

\subsection{F1000Research}

@@

This manuscript was developed entirely as an open project on GitHub. GitHub is
one of several hosting services for collaborative development using the Git
version control system (VCS).  Git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous
contributers to work asynchronously on the same project, often in parallel
branches, all of which can be effortlessly merged and version controlled.  While
Git is created primarily for software development, where the use of version
control systems is standard \cite{aru12}, it is ideal for academic research
since it provides a way to collaborate on every step of the manuscript
development process, from data manipulation and analysis to writing and
revision. For example, during the development of this manuscript, each author
would clone the project (\emph{i.e.} make a personal copy), modify it, and then
merge their changes into a master branch. This takes the preprint process to an
entirely new level, where the entire writing process is open from the beginning.

@@

Developing a preprint culture in biology will not solve all problems with the
current publication process. However, It might significantly reduce its negative
consequences. Preprint servers speed up publication and make no judgement on
pertinence and originality. Paradoxically, because peer-reviewing is conducted
\emph{after} this step in the process of scientific communication, it might
bring it closer to its fundamental objective. The role of peer-reviewing is to
judge the scientific quality of a study. The peer review process is the first
barrier against the fraudulent and poor quality science susceptible to impede
scientific progress. Preprints are not filtered, neither for their quality nor
their pertinence.  Technically, the difference between a preprint and a
traditional publication should be that the latter as the approval stamp by pairs
on its quality. The relevance of a study, its contribution to science, should
only be judged post-publication by many more readers than the typical two-four
anonymous reviewers. With a such a shift in the diffusion strategy, the role of
traditional journals and their editors would be to showcase scientific
discoveries for specialized readership. Such a process should improve even
further their diffusion, not impede it.

>>>>

Public preprint servers allow authors to make manuscripts publicly available
before, or in parallel to, submitting them to journals for traditional peer-
review. The rationale for preprint servers is fundamentally simple: to make
the results of research available to the scientific community as
soon as possible, instead of waiting until the peer-review process is fully
completed.  Sharing manuscripts using preprint servers has numerous advantages
including: 1) rapid dissemination of work-in-progress to a wider audience; 2)
immediate visibility of the research output for early-career scientists; 3)
improved peer review by encouraging feedback from the entire research
community; and 4) a fair and straightforward way to establish precedence.

Open preprint servers offer a great opportunity for open science, especially if
the community embraces the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work \cite{lom12}. These initiatives are
important to fully exploit the potential of open preprint servers. Posting
preprints online increases the community of available informal peer reviewers,
and uses the internet for its original community-building purposes.

Preprints began to gain popularity 20 years ago with the advent of arXiv, an
open preprint server widely used in physics and mathematics \cite{gin11}.
Preprints are also integral to the culture of other scientific fields.  Paul
Krugman noted that, in economics, the \emph{traditional model of submit, get
refereed, publish, and then people will read your work broke down a long time
ago. In fact, it had more or less fallen apart by the early 80s} \cite{kru12}.
In addition to a section in arXiv, economists have the RePEc (Research Papers in
Economics) initiative, which aims to create an archive of working papers,
manuscripts, and book chapters.

Despite the success of this approach in other fields, most manuscripts in
biology are not posted to preprint servers and are therefore not seen by more
than a handful of other scientists prior to publication. In this article, we: 1)
highlight the advantages of open preprint servers for both scientists and
publishers; 2) discuss the preprint policies of major publishers in biology; and
3) review the most popular preprint servers currently available.

@@

Public preprints can be crucial to early-career scientists. The delay before
publication is seldom compatible with the pressure to show an impressive
publication record when applying for a scholarship or a position. Increasing
the perceived value of pre-prints as close, or equal, to journal articles 
will allow young researchers to put their
research outcome in the open, and build a reputation for themselves through
the diffusion of their work without fear that this work will not be
recognized by grant or job committees. In this perspective, mixed publication
systems like F1000Research [\ref{f1000r}] should be encouraged as
they combine the best of both worlds: contributions are initially considered
preprints, but acquire full article status after positive reviews are
received.

Posting manuscripts as preprints also has the potential to result in higher
quality science by allowing prepublication feedback from a large pool of
reviewers. Prepublication reviews by a small network of colleagues are common in
the biological sciences and form an important part of the scientific process.
These ``friendly'' reviews increases the chances of errors being caught prior to
publication. Furthermore, the formal peer-review process as a whole is
critically over-loaded. As the number of active scientists increases and the
pressure to publish increases, it is becoming difficult for journals
to find reviewers \cite{hoc09}.  At the same time, rejection rates are high in
most journals \cite{aar08,roh09}, and when not invited to submit a revision,
authors must start the process over again at another journal. As a result,
initiatives to reduce time from submission to publication have emerged across
the scientific community. Rohr et al.  \cite{roh09} called for the recycling and
reuse of peer-reviews: by attaching previous reviews and detailed replies to a
new submission, both the editor and the referees can gauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. A widespread use of
preprint servers can achieve the same goal of reducing the time spent in review.
With a rich enough community of scientists depositing preprints, and commenting
on them, the process of an open pre-review can become widespread and will
overall increase the quality of first submissions \cite{hoc12}.

@@

Finally, public preprint servers offer a fair way to establish intellectual
priority by making the work available as soon as it is complete. Some
manuscripts will spend much more time than others in the review process and/or in
production after acceptance. This means that publication and
acceptance dates do not accurately characterize who came up with an idea
first. For this reason, mathematicians and physicists have embraced arXiv in
part to establish priority in a fair way \cite{gin11,cal12}.

@@

There are a number of reasons that biologists have not traditionally had a
culture of sharing preprints, many of which are based on common misconceptions
regarding the costs and benefits associated with the practice. For example, in contrast
to other fields there is a perception in biology that public preprints
make it easier to steal ideas. However, there is no evidence of this happening
in the numerous other fields that have adopted preprint servers, and since preprint
servers create a clear record of who had the idea first, and when, this
appears to be a largely unfounded concern. In other fields preprints serve the
opposite role, they allow straightforward establishment of precedence, letting
research lay claim to an idea thus preventing it from being ``stolen'' \cite{gin11}

Another major concern is that if a researcher posts a public preprint that they
will not be allowed to submit their paper to their journal of choice.  This is
based on a certain interpretation of the Ingelfinger rule: scientists should not
publish the same manuscript twice \cite{alt96}. A preprint is simply a document
that allows ideas to spread and be discussed, it is not yet formally validated
by the peer-review system. This is why the majority of publishers do not see
arXiv and similar services as a violation of the Ingelfinger rule. Almost all of
the major publishers in biology are preprint-friendly, including: Nature
Publishing Group, PLOS, BMC, PNAS, Elsevier, and Springer \ref{table:policies}.
The Ecological Society of America and the Genetics Society of America recently
changed their policies to allow public preprints. \emph{Nature}
responded to the rumour that they refused manuscripts submitted to arXiv by
saying that ``\emph{Nature} never wishes to stand in the way of communication
between researchers. We seek rather to add value for authors and the community
at large in our peer review, selection and editing'' \cite{nat05}.  Still, a few
journals adopt a ``by default'' hostile attitude towards preprints, mostly due
to the lack of clear policy of the publishers, or perhaps because a preprint
culture has not developed in biology and the practice is still considered
unusual. As an example, Wiley-Blackwell, which publishes some of the leading
journals in biology, has no official policy on the matter.

@@

One of the advantages of figshare over arXiv for biologists is that is it not
limited to quantitative sciences. arXiv.org has sections on quantitative biology
but might not be appropriate for non-quantitative work. With its flexible
approach to preprints, figshare offers an alternative to arXiv for empirical
biologists. Furthermore, by allowing all types of content, figshare also
provides an archive for early results (\emph{e.g.}: figures, lab presentations).

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new open access
publisher that combines both a preprint server and a peer reviewed journal.  It
is focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a place in biology. Like figshare this is
an advantage relative to arXiv for biologists doing non-quantitative work.  Also
like figshare, PeerJ allows commenting on posted preprints, improving the
potential for pre-publication dialog. In addition, preprints can be made private
if the authors choose, and shared only with selected colleagues. While this
reduces many of the benefits of preprints described above, it may allow some
researchers who would not otherwise post preprints to begin exploring the
possibility in a manner appropriate to their current circumstances.

In contrast to other preprint servers, users cannot post unlimited public
preprints for free. One preprint per year can be posted for free, with a onetime
(\emph{i.e.} lifetime) fee of 99 dollars required to allow an author to post unlimited
public preprints. The preprint server is not tied to the journal, so preprints
can be posted regardless of where they will eventually be submitted for
publication.

\subsection{F1000Research}\label{f1000r}

@@

This manuscript was developed entirely as an open project on GitHub. GitHub is
one of several hosting services for collaborative development using the Git
version control system (VCS).  Git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous
contributers to work asynchronously on the same project, often in parallel
branches, all of which can be effortlessly merged and version controlled.  While
Git is created primarily for software development, where the use of version
control systems is standard \cite{aru12}, it is ideal for academic research
since it provides a way to collaborate on every step of the manuscript
development process, from data manipulation and analysis to writing and
revision. For example, during the development of this manuscript, each author
would clone the project (\emph{i.e.} make a personal copy), modify it, and then
merge their changes into a master branch. This takes the preprint process to an
new level, where the entire writing process is open from the beginning.

@@

Developing a preprint culture in biology will not solve all problems with the
current publication process. However, It might significantly reduce its negative
consequences. Preprint servers speed up publication and make no judgement on
pertinence and originality. Paradoxically, because peer-reviewing is conducted
\emph{after} the manuscript is made public, preprint servers might bring it
closer to its fundamental objective. The role of peer-reviewing is to judge the
scientific quality of a study. It is the first barrier against the fraudulent
and poor quality science susceptible to impede scientific progress. Preprints
are not filtered, neither for their quality nor their pertinence.  Technically,
the difference between a preprint and a traditional publication should be that
the latter as the approval stamp by pairs on its quality. The relevance of a
study, its contribution to science, should only be judged post-publication by
many more readers than the typical two-four anonymous reviewers. With a such a
shift in the diffusion strategy, the role of traditional journals and their
editors would be to showcase scientific discoveries for specialized readership.
Such a process should improve even further their diffusion, not impede it.
